### 2019-10-28 paper.tex
Initlialize section names.

\documentclass[twocolumn]{article}
\usepackage{cite}
\usepackage[margin=1in]{geometry}
\usepackage{url}
\title{OmniVisor - OmniScale Hypervisor for Disaggregated Architecture}
\author{Yu-Ju Huang}
\date{Cornell University}
\begin{document}
\maketitle

\begin{abstract}
Put your abstract here.
\end{abstract}

% This is a comment.
Welcome to your first \LaTeX paper!

\section{Introduction}

\section{Background}

\begin{itemize}
\item \texttt{pdflatex latex}
\item \texttt{pdflatex latex}
\item \texttt{bibtex latex}
\item \texttt{pdflatex latex}
\item \texttt{pdflatex latex}
\end{itemize}

\section{Related Work}

Perhaps you're looking to cite Dijkstra's paper on
THE\cite{DBLP:journals/cacm/Dijkstra68}.  Just put the citation in the bib file of the same name as your main file.

\section{Design}
\section{Implementation}
\section{Evaluation}

%latex here should be the name of your bibtex file minus '.bib'
\bibliography{paper}{}
\bibliographystyle{plain}

\end{document}